---
title: "Project 5 - Breast Cancer Classification"
author: "Group 2"
date: "4/25/2018"
output:
  html_notebook: default
  pdf_document: default
---
## Table of Contents  
[Introduction](#Intro)  
[Step 0: Load Packages & Specify Directories](#Step0)  
[Step 1: Load and Process Data](#Step1)  
[Step 2: Feature Selection](#Step2)  
[Step 3: Implement Algorithm](#Step3)  
[Step 4: Evaluation](#Step3)  

### Introduction {#Intro}   

### Step 0: Load Packages & Specify Directories {#Step0}  
```{r}
# Packages that will be used
packages.used <- c("PerformanceAnalytics", "ggplot2")
# Check packages that need to be installed
packages.needed <- setdiff(packages.used, 
                           intersect(installed.packages()[,1], 
                                     packages.used))
# Install additional packages
if(length(packages.needed)>0){
  install.packages(packages.needed, dependencies = TRUE,
                   repos='http://cran.us.r-project.org')
}
# Load libraries  
library("PerformanceAnalytics")
library("ggplot2")

# Set working directory to the doc folder 
#setwd("~/GitHub/Spring2018-Project5-grp_2/doc")
```

### Step 1: Load and Process Data {#Step1}  
#### Load Data
```{r}
df <- read.csv("../data/data.csv", header = TRUE, stringsAsFactors = FALSE)
```
#### Explore Data {.tabset}
**Attribute Information:**  
1) ID number  
2) Diagnosis (M = malignant, B = benign)   
3-32) Ten real-valued features described as follows are computed for each cell nucleus:  

a) radius (mean of distances from center to points on the perimeter)  
b) texture (standard deviation of gray-scale values)  
c) perimeter  
d) area  
e) smoothness (local variation in radius lengths)  
f) compactness (perimeter^2 / area - 1.0)  
g) concavity (severity of concave portions of the contour)  
h) concave points (number of concave portions of the contour)   
i) symmetry  
j) fractal dimension ("coastline approximation" - 1)   

The mean, standard error and "worst" or largest (mean of the three largest values) of these features were computed for each image, resulting in 30 features. For instance, field 3 is Mean Radius, field 13 is Radius SE, field 23 is Worst Radius.   

33) All entries are NA's   

##### Head  
```{r}
head(df)
```

##### Structure  
```{r}
str(df)
```
##### Summary
```{r}
summary(df)
```

#### Process Data  
```{r}
# Delete last column from dataset as its entries are all NA's
df <- df[,-1]

# Factorize the diagnosis attribute
df$diagnosis <- factor(ifelse(df$diagnosis=="B","Benign","Malignant"))

# Split entire data into 80% train set and 20% test set
set.seed(123)
index <- sample(1:nrow(df),0.8*nrow(df))
df.train <- df[index,]
df.test <- df[-index,]

# Check proportion of diagnosis (Benign/Malignant) in train/test sets
#prop.table(table(df.train$diagnosis))
#prop.table(table(df.test$diagnosis))
```

### Step 2: Feature Selection {#Step2}
 

### Step 3: Implement Algorithm  {#Step3}  


#### algorithms {.tabset}  
##### Random Forest    
##### Logistic Regression  
##### GBM  
##### XGBoost  
##### AdaBoost  
##### SVM    

### Step 4: Evaluation  {#Step4} 